---
title: "Exploratory_analysis_insurance"
author: "Jiayi Shen"
date: "12/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Reference to the association between insurance type and cancer mortality
- [Trends in Cancer Survival by Health Insurance Status in California From 1997 to 2014., 2018](https://www.ncbi.nlm.nih.gov/pubmed/29192307)   
Relative to privately insured patients, cancer-specific mortality was higher in uninsured patients for all cancers except prostate, and disparities were largest from 2009 to 2014 for breast (HR, 1.72; 95% CI, 1.45-2.03), lung (men: HR, 1.18; 95% CI, 1.06-1.31 and women: HR, 1.32; 95% CI, 1.15-1.50), and colorectal cancer (women: HR, 1.30; 95% CI, 1.05-1.62). Mortality was also higher for patients with other public insurance for all cancers except lung, and disparities were largest from 2009 to 2014 for breast (HR, 1.25; 95% CI, 1.17-1.34), prostate (HR, 1.17; 95% CI, 1.04-1.31), and colorectal cancer (men: HR, 1.16; 95% CI, 1.08-1.23 and women: HR, 1.11; 95% CI, 1.03-1.20).  

- [The Impact of Health Insurance Status on the Survival of Patients With Head and Neck Cancer, published on 2010](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3085979/)
Patients with Medicaid/uninsured (hazard ratio [HR], 1.50; 95% confidence interval [CI], 1.07-2.11) and Medicare disability (HR, 1.69; 95% CI, 1.16-2.48) had significantly lower overall survival compared with patients with private insurance

- [Factors That Contributed to Black-White Disparities in Survival Among Nonelderly Women With Breast Cancer Between 2004 and 2013, published on 2017](http://ascopubs.org/doi/abs/10.1200/jco.2017.73.7932)
Matching by insurance explained one third of the excess risk of death;  
the HR for death in black versus white women is 1.54 (95% CI, 1.47 to 1.62) when matched for insurance.

```{r load data}
can_reg = read.csv("Cancer_Registry.csv")
can_reg_insurance = select(can_reg, PctPrivateCoverage:PctPublicCoverageAlone)

#return rows with NAs with respect to insurance variables.
dim(can_reg_insurance[which(!complete.cases(can_reg_insurance)),])
```

- 609 missing values in `PctPrivateCoverageAlone`.

## Distribution of variables 
```{r}
library(GGally)
par(mfrow = c(3, 2))
hist(can_reg$PctPrivateCoverage)
hist(can_reg$PctPrivateCoverageAlone)
hist(can_reg$PctEmpPrivCoverage)
hist(can_reg$PctPublicCoverage)
hist(can_reg$PctPublicCoverageAlone)
```

- histograms are pretty much bell-shaped. 
- more counties with high pct of private coverage, low pct of public coverage. 



## reference category
```{r}
# try make a reference group (i.e. uninsuranced??)
can_reg_insurance %>% 
  mutate(no_insurance = 100 - 0.5*(PctPrivateCoverage + PctPrivateCoverageAlone + PctPublicCoverage + PctPublicCoverageAlone)) %>% pull(no_insurance)
```

Questions:
- do we need to do stratified analysis?? (by age, gender and other factor of interest)


