---
title: "JX-backward"
author: "Jingyu Xu"
date: "2018/12/9"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
```{r}
#load and exculde redundant variable
can_reg = read.csv(file = "./Cancer_Registry.csv")
can_reg_omit = can_reg %>% 
  janitor::clean_names() %>% 
  select(-pct_some_col18_24, -pct_no_hs18_24, -pct_hs18_24, -pct_bach_deg18_24, -avg_ann_count, -avg_deaths_per_year, -med_income, -pop_est2015, -median_age_male, -median_age_female, -percent_married, -pct_employed16_over, -pct_private_coverage, -pct_private_coverage_alone, -pct_emp_priv_coverage, -pct_public_coverage, -pct_black, -pct_asian, -pct_other_race) %>% 
  na.omit()%>%
  filter(median_age < 100)%>%
  select(geography, target_death_rate, binned_inc, everything())
```
```{r}
#description of raw dataset and cleaned dataset
summary(can_reg)
summary(can_reg_omit)
```
##Backward elimination (AIC based)
```{r}
#backward elimination
 can_reg_back = can_reg_omit%>%
  select(-geography)
initial_fit = lm(target_death_rate~., data = can_reg_back)
step(initial_fit, direction = "backward")
```
After backward elimination, the final model is : 
target_death_rate ~ binned_inc + incidence_rate + 
    median_age + pct_hs25_over + pct_bach_deg25_over + pct_unemployed16_over + 
    pct_public_coverage_alone + pct_married_households + birth_rate
    
There are 9 variables in the model.
##Check the distribution of variables
```{r}
##see the distribution of each variables
ggplot(gather(select(can_reg_back, -binned_inc)), aes(value)) + 
    geom_histogram(bins = 8) + 
    facet_wrap(~key, scales = 'free_x')
```
pct_white, poverty_percent and study_per_cap are a little skewed, other variables are nearly symmetric
##Model diagnosis
```{r}
##summary of final model
back_model = lm(formula = target_death_rate ~ binned_inc + incidence_rate + 
    median_age + pct_hs25_over + pct_bach_deg25_over + pct_unemployed16_over + 
    pct_public_coverage_alone + pct_married_households + birth_rate, data = can_reg_back)
#check model assumption
summary(back_model)
par(mar=c(2,2,1,1))
plot(back_model)
#check multicollinearity
faraway::vif(back_model)%>%knitr::kable()
```
*  From residual and scale location figure, I found the points distributed evenly around the horizontal line, indicating the constant variance
*  From QQ plot, we can observe a tail deviating from normal, which doubts the normality of residuals.
* From Residual and Leverage plot, we observe the potential outlier: 1211, 1355 and 279
* The global F-statistic is 174.1,indicating the linear relationship between the outcome and the covariates is significant. However,there are still covariates that are insignifficant and the adjusted R square is low (0.4938).
*VIF: No variables has VIF over 5, indicating they don't have much overlapped information.
*Further modificaton may be needed(The adjusted R square went higher, but still requires improving)
##Model validation
```{r}
library(caret)
#Use 10-fold validation and create the training sets
data_train<-trainControl(method="cv", number=10)
# Fit the 4-variables model that we discussed in previous lectures
model_caret<-train(target_death_rate ~ binned_inc + incidence_rate + 
    median_age + pct_hs25_over + pct_bach_deg25_over + pct_unemployed16_over + 
    pct_public_coverage_alone + pct_married_households + birth_rate,
                   data=can_reg_omit,
                   trControl=data_train,
                   method='lm',
                   na.action=na.pass)
  
# Model predictions using 4 parts of the data fro training 
model_caret
# Model coefficients
model_caret$finalModel
# Examine model prediction for each fold
fold =model_caret$resample%>%
  as.tibble%>%
  mutate(MSE = RMSE^2)
# look at the summary of MSE
summary(fold$MSE)
```