---
title: "Single Cancer Analysis and Figures"
author: "James Dalgleish"
date: "December 14, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = T, warning = F, message = F)
library(tidyverse)
library(corrplot)
library(glmnet)
library(plotmo)
library(doParallel)
library(ggtern)
library(selectiveInference)
```
```{r lasso_with_redundant_removed, fig.width=20, fig.height=20}
#saveRDS(can_reg_joined,"can_reg_joined.rds")
# saveRDS(can_reg_state,"can_reg_state.rds")
# saveRDS(can_reg,"can_reg.rds")
#saveRDS(can_reg,"can_reg.rds")
grid <- 10^seq(5,-2, length=100)
can_reg_joined<-readRDS("can_reg_joined.rds")
can_reg_state<-readRDS("can_reg_state.rds")
cancer_names <- list.files("can_inc") %>% gsub(x = ., pattern = ".csv", replacement = "")
can_filt3 <- can_reg_joined %>% 
  janitor::clean_names() %>% 
    mutate(binned_inc = as.factor(binned_inc)) %>% 
  mutate(binned_inc = fct_reorder(.f = binned_inc,.x = med_income)) 
#add_top_can
can_filt3$top_can <- apply(can_filt3[,cancer_names],1,which.max) %>% map(.x = ., .f = ~ifelse(length(.)==0,"none",cancer_names[.])) %>% unlist() %>% as.factor()
#apply(can_filt[,cancer_names],1,which.max) %>% unlist() %>% table() %>% names(.) %>% as.numeric() %>% cancer_names[.]
can_filt3 <- can_filt3 %>% 
  #transmute(avg_household_size = log(avg_household_size)) %>% 
#  mutate()
#  .[c(-1009,-2221,-2090,-1008,-2087,-2218, -1007,-1006,-1005,-1004,-1003, -1002),] %>% 
dplyr::select( -pop_est2015, -med_income, -avg_ann_count, -avg_deaths_per_year, -median_age, -median_age_male,        -pct_no_hs18_24, -pct_hs18_24, -pct_bach_deg18_24, -pct_some_col18_24,  -percent_married, -pct_employed16_over, -pct_private_coverage, -pct_private_coverage_alone, -pct_emp_priv_coverage, -pct_public_coverage, -pct_black, -pct_asian, -pct_other_race, -state_abb,  -county, -state, -geography,  -binned_inc, -uterine, -bladder, -ovary, -brca_cis, -cervix, -colorectal, -melanoma, -ovary, -prostate, -brca, -study_per_cap) #%>%
  #mutate(study_per_cap = study_per_cap^(-1)) %>% 
  #na.omit()
#  mutate(study_per_cap = atan(study_per_cap))

X = as.matrix(can_filt3 %>% dplyr::select(-target_death_rate) %>% sapply(as.numeric) %>% na.omit())
Y = na.omit(can_filt3)$target_death_rate
set.seed(1)
train<-sample(1:nrow(X),nrow(X)/2)
test<-(-train)
Y.test<-Y[test]
set.seed(2)
doParallel::registerDoParallel()
cv.out<-cv.glmnet(X[train,],Y[train], alpha=0.5,nfolds = 10, parallel = TRUE, lambda = grid)
plot(cv.out)
best.lambda<-cv.out$lambda.min
lasso.pred <- predict(cv.out,s=best.lambda,newx=X[test,])
final_lasso_with_inc_reduced_single_can_MSE <-mean((lasso.pred-Y.test)^2) #MSE
lasso3<-glmnet(X, Y, alpha=1, lambda=best.lambda)
res_lasso <- coef(lasso3) 
res_lasso %>% as.matrix() %>% as.data.frame() %>% add_rownames() %>% as.tibble() %>% 
  mutate(abs_lasso = abs(s0)) %>% 
  arrange(-abs_lasso) %>% knitr::kable()
final_lasso_with_inc_reduced_single_can<-lasso3
final_lasso_with_inc_reduced_single_can$dev.ratio
final_lasso_range<-glmnet(X, Y, alpha=1, lambda=grid)
plotmo::plotres(final_lasso_range, info = T,which = 1:4) 
 # res_lasso %>% as.matrix() %>% as.data.frame() %>% add_rownames() %>% as.tibble() %>% 
 #  mutate(abs_lasso = abs(s0)) %>% 
 #  arrange(-abs_lasso) %>% View()

```

```{r corrplot}
corrplot(cor(X))
```

We have now created another model with dev ratio of `r final_lasso_with_inc_reduced_single_can$dev.ratio` and MSE = `r final_lasso_with_inc_reduced_single_can_MSE`. The model features are not quite so stunning, but it is simpler, easier to interpret, and incorporates more data. We'll find that this model doesn't put brca at the top. In fact, it's almost completely shrunken from the model entirely. There could be a lot of reasons for this, but one might be that the 5 year survival rate for breast cancer is nearly 90%. This could be attributable to treatment regimens, but other cancers may be easier to predict death rate.
```{r spacing_test}

lar_full<-selectiveInference::lar(x = X,y = Y)
spacing_test<-selectiveInference::larInf(lar_full) #this is a better test that actually gives inference of the individual predictor's correlation.
spacing_test$predictor <- colnames(X)[spacing_test$var]
#class(spacing_test_table)<-"list"
ci_tibble <- spacing_test$ci %>% as.tibble()
colnames(ci_tibble)<-c("upper_CI","LowerCI")
covtest_spacing_pval <- tibble(predictor = colnames(X)[spacing_test$vars],spacing.p.value = spacing_test$pv.spacing, covtest.pval = spacing_test$pv.covtest) %>%
  bind_cols(ci_tibble) %>% 
  mutate(significance = ifelse(spacing.p.value < 0.05, "*","")) %>% 
  mutate(adj.spacing.pvalue = p.adjust(spacing.p.value)) %>% 
  mutate(significance_adj = ifelse(adj.spacing.pvalue < 0.05, "*","")) %>%
  select(covtest.pval,spacing.p.value,significance,adj.spacing.pvalue,significance_adj,everything()) %>% 
  dplyr::arrange(spacing.p.value)
#nrow(X)
#spacing_test_table %>% coef.lar()
#length(Y)
covtest_spacing_pval %>% knitr::kable()
#covtest_spacing_pval %>% View
```
Using a test suggested by Tshibirani in his text, (statistical learning with sparsity, the lasso and its generalizations, p.161), we find that the lung-bronchial cancer incidence rate, the raw incidence rate for all cancers, and both college education and high school education status have a nonzero coefficient in the least angle regression model. The text suggests that this test is superior to the covariance test in that the entire model is not tested at each knot, but rather "whether the partial correlation of the given predictor entered in that step [knot] is zero, adjusting for other variables currently in the model" (p.161). As such, it does not simply test for the lasso path, but for the individual contribution of each predictor to the model. As we might gather, this agrees completely with the covariance test results and with the confidence intervals that do not contain zero in the least angle regression estimate confidence interval. It is often said that p-values are not sharp cutoffs and the adjustment procedure here has been very conservative. 
```{r missing}
#there is some missing data, but only in the brca incidence.
Amelia::missmap(can_filt3)
```

```{r averages_by_region}
#which region has the highest BRCA incidence?
can_filt3 %>% 
  na.omit() %>% 
  group_by(state_region) %>% 
  # summarize(mean_brca = mean(brca),
  #           mean_pct_public_coverage_alone = mean(pct_public_coverage_alone))
  summarise_all("mean") %>% 
  gather(variable,region_mean, -state_region) %>% 
  ggplot(aes(y=variable,x=region_mean, color=state_region)) +
  geom_point() + theme(axis.text.x = element_text(angle = 90, hjust = 1))
  # skimr::skim() %>% 
  # arrange(variable, region)
#does brca play a bigger role there?
#log transform everything that needs it.
```

```{r}
can_filt3 %>% 
  gather(variable,value,-state_region) %>% 
  mutate(value = as.numeric(as.character(value))) %>% 
  mutate(value = heatmaply::percentize(value)) %>% 
  # group_by(variable) %>% 
  # summarize(transformed_value = heatmaply::percentize(value)) %>% 
#  filter(variable != "study_per_cap") %>% 
     na.omit() %>% 
  # group_by(state_region) %>% 
  # summarize(mean_brca = mean(brca),
  #           mean_pct_public_coverage_alone = mean(pct_public_coverage_alone))
  ggplot(aes(x=variable, y = value ,  color = state_region)) + #group=state_region
  geom_boxplot() + theme(axis.text.x = element_text(angle = 90, hjust = 1))

  # summarise_all("mean") %>% f
  # gather(variable,region_mean, -state_region) %>% 
  # ggplot(aes(y=variable,x=region_mean, color=state_region)) +
  # geom_point() + theme(axis.text.x = element_text(angle = 90, hjust = 1))
can_reg_state %>%
  janitor::clean_names() %>% 
  dplyr::select(state,county,incidence_rate, everything()) %>% 
  dplyr::arrange(-incidence_rate)
#cancers in the south tend to
```

```{r}
can_filt3 %>% 
  mutate(high_poverty = poverty_percent > can_filt3$poverty_percent %>% summary() %>% .["3rd Qu."]) %>% 
  gather(variable,value,-high_poverty) %>% 
  mutate(value = as.numeric(as.character(value))) %>% 
  mutate(value = heatmaply::percentize(value)) %>% 

#  filter(variable != "study_per_cap") %>% 
     na.omit() %>% 
  # group_by(state_region) %>% 
  # summarize(mean_brca = mean(brca),
  #           mean_pct_public_coverage_alone = mean(pct_public_coverage_alone))
  ggplot(aes(x=variable, y = value ,  color = high_poverty)) + #group=state_region
  geom_boxplot() + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
```{r poverty_boxplot}
can_filt3 %>% 
  mutate(racial_diversity = pct_white < can_filt3$poverty_percent %>% summary() %>% .["1st Qu."]) %>% 
  gather(variable,value,-racial_diversity) %>% 
  mutate(value = as.numeric(as.character(value))) %>% 
  mutate(value = heatmaply::percentize(value)) %>% 

#  filter(variable != "study_per_cap") %>% 
     na.omit() %>% 
  # group_by(state_region) %>% 
  # summarize(mean_brca = mean(brca),
  #           mean_pct_public_coverage_alone = mean(pct_public_coverage_alone))
  ggplot(aes(x=variable, y = value ,  color = racial_diversity)) + #group=state_region
  geom_boxplot() + theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

color ternary plot-- 4 variables, 5 with a shape.
https://rdrr.io/cran/ggtern/
poverty, educaction, etc. many variables tend to go together, especially with lung cancer.
smoking problably is a good way to go as well.
access to treatments.
percent public coverage alone-- show it's relationship to death rate and incidence rate
```{r ternary, figure.width = 10 , figure.height = 10}
#could do a ternary for each region (facet)
#https://rpubs.com/KDVdecisions/triadtutorial1
# can_filt3 %>% 
#   mutate(pct_nonwhite = 1-pct_white) %>%
#   ggplot(data=.) %>% 

ggtern_plot <-  ggtern::ggtern(data=na.omit(can_filt3), mapping= aes(x=fmsb::percentile(poverty_percent),y=fmsb::percentile(target_death_rate),z=fmsb::percentile(1-pct_bach_deg25_over),facet=state_region, color = fmsb::percentile(pct_public_coverage_alone)), geom='polygon') +
  stat_density_tern(aes(fill=..level.., alpha=..level..),geom='polygon') +
  scale_fill_gradient2(high = "red") +         
   geom_point() + facet_wrap(~state_region,ncol=2) + scale_color_viridis_c() + geom_density_tern() + theme_showarrows() +
  guides(fill = "none", alpha = "none") +
  labs (x = "Poverty", y = "Death Rate", z = "% Bachelor's Degree", color = "% Public Insurance Coverage Alone") +  theme(legend.position = "bottom") + Tlab(label = "", labelarrow = "Death Rate") + Llab(label = "", labelarrow = "Poverty") + Rlab(label = "", labelarrow = "% witout Bach")
ggtern_plot
#par(mfrow)
# can_filt3[can_filt3$state_region=="South",c("poverty_percent","target_death_rate","pct_bach_deg25_over")] %>% transmute(poverty_percent = poverty_percent, target_death_rate, pct_without_deg = 100-pct_bach_deg25_over) %>% cor(use = "pairwise.complete.obs", method="kendall") %>%  corrplot(order="hclust", method = "number")
# can_filt3[can_filt3$state_region=="South",c("poverty_percent","target_death_rate","pct_bach_deg25_over")] %>% cor(use = "pairwise.complete.obs", method="kendall") %>%  corrplot(order="hclust", method = "number")
# can_filt3[can_filt3$state_region=="Northeast",c("poverty_percent","target_death_rate","pct_bach_deg25_over")] %>% cor(use = "pairwise.complete.obs", method="kendall") %>%  corrplot(order="hclust", method = "number")
# can_filt3[can_filt3$state_region=="North Central",c("poverty_percent","target_death_rate","pct_bach_deg25_over")] %>% cor(use = "pairwise.complete.obs", method="kendall") %>%  corrplot(order="hclust", method = "number")
# can_filt3[can_filt3$state_region=="West",c("poverty_percent","target_death_rate","pct_bach_deg25_over")] %>% cor(use = "pairwise.complete.obs", method="kendall") %>%  corrplot(order="hclust", method = "number")
foreach::foreach(reg=levels(can_filt3$state_region)) %do%
{
  can_filt3[can_filt3$state_region==reg,c("poverty_percent","target_death_rate","pct_bach_deg25_over","pct_public_coverage_alone")] %>% transmute(poverty_percent = poverty_percent, target_death_rate,pct_public_coverage_alone = pct_public_coverage_alone, pct_without_deg = 100-pct_bach_deg25_over) %>% cor(use = "pairwise.complete.obs", method="pearson") %>%  corrplot( method = "number", title=paste0(reg,"region"))
}
```

It's compelling that the relationship between poverty and death rate is strongest in certain regions, namely in the west. A more indepth view of states within the west state areas this may provide additional insight.  We'll add back in the state and county data to see which states/counties may be responsible for the incredibly weak link between poverty and death rate in the west. Ternary plots are uncommon, but a powerful way to display complex relationships of three independent variables (http://science.sciencemag.org/content/sci/215/4537/1191.full.pdf).
Figure 4. Ternary plot suggests that the 3-way relationship between education, poverty, insurance coverage differs as the polygon placement and  shape differs, suggesting that numerical relationships between these three variabes is not equally strong in all pairwise relationships. Notably, we also find a large central peak in the south plot that shows that each of all three variables are related to each other (percentages of poverty, of indiviudals without a bachelor's degree, and percentile death rate). The high amount brightly colored points suggests counties with a high percentage of indiviudals with only public coverage. In short, all of the variables are related well in the  south, but not so in the west (which contains lower income areas that are highly educated and with the lowest cancer areas in the country).


```{r}
can_filt4 <- can_reg_joined %>% 
  janitor::clean_names() %>% 
    mutate(binned_inc = as.factor(binned_inc)) %>% 
  mutate(binned_inc = fct_reorder(.f = binned_inc,.x = med_income)) 
can_filt4$top_can <- apply(can_filt4[,cancer_names],1,which.max) %>% map(.x = ., .f = ~ifelse(length(.)==0,"none",cancer_names[.])) %>% unlist() %>% as.factor()
#apply(can_filt[,cancer_names],1,which.max) %>% unlist() %>% table() %>% names(.) %>% as.numeric() %>% cancer_names[.]
can_filt4 <- can_filt4 %>% 
  #transmute(avg_household_size = log(avg_household_size)) %>% 
#  mutate()
#  .[c(-1009,-2221,-2090,-1008,-2087,-2218, -1007,-1006,-1005,-1004,-1003, -1002),] %>% 
dplyr::select( -pop_est2015,  -avg_ann_count, -avg_deaths_per_year, -median_age, -median_age_male,        -pct_no_hs18_24, -pct_hs18_24, -pct_bach_deg18_24, -pct_some_col18_24,  -percent_married, -pct_employed16_over, -pct_private_coverage, -pct_private_coverage_alone, -pct_emp_priv_coverage, -pct_public_coverage, -pct_black, -pct_asian, -pct_other_race, -state_abb,   -geography,  -binned_inc, -uterine, -bladder, -ovary, -brca_cis, -cervix, -colorectal, -melanoma, -ovary, -prostate, -brca, -study_per_cap)

```
```{r}
can_filt4 %>% 
  filter(state_region == "West") %>% 
  group_by(state) %>% 
  summarize(dr_pov_cor = cor(target_death_rate,poverty_percent),
            abs_dr_pov_cor = abs(dr_pov_cor)) %>% 
  arrange(abs_dr_pov_cor)
```
We'll note that, on the whole, in the western states, poverty percentage has a very weak linear relationship with cancer death rate. Utah and Oregon are the exceptions. Alaska and Hawaii are the exceptions. These states tend to have lower smoking rates generally(https://www.cdc.gov/tobacco/data_statistics/fact_sheets/adult_data/cig_smoking/index.htm), although honestly the reasons behinid this lack of poverty disparity is likely has many influencing factors. It may be that poverty doesn't adequately describe some of these areas and so we'll look at median income's relationship (even though including it would have created issues of redundant measures or multicollinearity).
```{r}
can_filt4 %>% 
  filter(state_region == "West") %>% 
  group_by(state) %>% 
  summarize(dr_inc_cor = cor(target_death_rate,med_income),
            abs_dr_inc_cor = abs(dr_inc_cor)) %>% 
  arrange(abs_dr_inc_cor)
lm(data=can_filt4 %>% 
  filter(state_region == "West") %>% select(-state_region, -county, -state),target_death_rate ~ .) %>% 
  broom::tidy() %>% 
  mutate(p.value = p.adjust(p.value)) %>% 
  arrange(p.value)
```
We using the above model with adjusted p-values, we find that the estimate for median income in this region is near zero. Education remains as the highest socioeconomic predictor.

```{r}
can_filt4_west <- can_filt4 %>% 
    filter(state_region == "West") 
X = as.matrix(can_filt4_west %>% dplyr::select(-target_death_rate, -state, -county, -state_region) %>% sapply(as.numeric) %>% na.omit())
Y = na.omit(can_filt4_west)$target_death_rate
set.seed(1)
train<-sample(1:nrow(X),nrow(X)/2)
test<-(-train)
Y.test<-Y[test]
set.seed(2)
doParallel::registerDoParallel()
cv.out<-cv.glmnet(X[train,],Y[train], alpha=0.5,nfolds = 10, parallel = TRUE, lambda = grid)
plot(cv.out)
best.lambda<-cv.out$lambda.min
lasso.pred <- predict(cv.out,s=best.lambda,newx=X[test,])
final_lasso_west_can_MSE <-mean((lasso.pred-Y.test)^2) #MSE
lasso3<-glmnet(X, Y, alpha=1, lambda=best.lambda)
res_lasso <- coef(lasso3) 
res_lasso %>% as.matrix() %>% as.data.frame() %>% add_rownames() %>% as.tibble() %>% 
  mutate(abs_lasso = abs(s0)) %>% 
  arrange(-abs_lasso) %>% knitr::kable()
final_lasso_with_inc_reduced_single_can<-lasso3
final_lasso_with_inc_reduced_single_can$dev.ratio
final_lasso_range<-glmnet(X, Y, alpha=1, lambda=grid)
plotmo::plotres(final_lasso_range, info = T,which = 1:4) 

```
After constructing a subset lasso model, we find that the number of factors decreases substantially, leaving education as the largest mangitude predictor, with married households and unemployment remaining with the incidence rates. What's interesting is the percentage of married households is important. This is somewhat unexpected.

```{r west_sig}
spacingTest<-function(X,Y)
{lar_full<-selectiveInference::lar(x = X,y = Y)
spacing_test<-selectiveInference::larInf(lar_full) #this is a better test that actually gives inference of the individual predictor's correlation.
spacing_test$predictor <- colnames(X)[spacing_test$var]
#class(spacing_test_table)<-"list"
ci_tibble <- spacing_test$ci %>% as.tibble()
colnames(ci_tibble)<-c("lower_CI","upper_CI")
covtest_spacing_pval <- tibble(predictor = colnames(X)[spacing_test$vars],spacing.p.value = spacing_test$pv.spacing, covtest.pval = spacing_test$pv.covtest) %>%
  bind_cols(ci_tibble) %>% 
  mutate(ci_midpoint = (upper_CI+lower_CI)/2) %>% 
  mutate(significance = ifelse(spacing.p.value < 0.05, "*","")) %>% 
  mutate(adj.spacing.pvalue = p.adjust(spacing.p.value)) %>% 
  mutate(significance_adj = ifelse(adj.spacing.pvalue < 0.05, "*","")) %>%
  select(covtest.pval,spacing.p.value,significance,adj.spacing.pvalue,significance_adj,everything()) %>% 
  dplyr::arrange(spacing.p.value)
}#nrow(X)
#spacing_test_table %>% coef.lar()
#length(Y)
spacingTest(X,Y) %>% knitr::kable()

```
The spacing test is even more stringent way to determine what variables are most important in the model than simply performing lasso and provides p-values. From this, we'll find that indeed married households play a role in reducing cancer death rate, at least in the west. We can say that the coefficient in the model is nonzero and the sign is negative, indicating an inverse relationship, in line with raw pearson correlation of $R=$`r cor(can_filt4_west$target_death_rate,can_filt4_west$pct_married_households)` between these variables in the western united states. An article published last year in *SSM population health* suggests several reasons for this (higher earning or higher education spouses tend to be beneficial for survival rates, https://doi.org/10.1016/j.ssmph.2016.12.007). The article is right in line with our model, which puts education as a highly important variable.
```{r marriage_rates_by_region}
can_filt4 %>% 
  filter(state_region == "West") %>% 
  group_by(state,county) %>% 
  # summarize(marriage_pct = mean(pct_married_households),
  #           marriage_pct = mean(pct_married_households)) %>% 
  select(state,county,pct_married_households,pct_bach_deg25_over,pct_hs25_over,med_income, everything()) %>% 
  arrange(-pct_married_households)
```
We look at the top counties by percent married and find that these locations tend to have a higher education than the regional average of `r mean(can_filt4$pct_married_households)` and national average of `r mean(can_filt4_west$pct_married_households)` percent. They are more educated than the national average of `r mean(can_filt4_west$pct_bach_deg25_over)`% individuals with a bachelor's degree and `r mean(can_filt4_west$pct_hs25_over)` with a high school education above 25. They also have lower death rates and ages than average as well.


We'll also note that the relationship between education, poverty, cancer death rate, and insurance has interesting regional trends.

```{r}
#excluding from report. filter(str_detect(string = estimate,":")) %>%
lm(data=can_reg_joined, target_death_rate ~ study_per_cap*state) %>% broom::tidy() %>%  filter(p.value < 0.05) %>%  arrange(estimate)
```

```{r}
#hit florida
#dichotonmize clinical trials
#2 stage lasso to boost dev ratio (cite book and 2007 paper).
can_reg_state[1000:1010,]
can_filt_florida <- can_reg_joined %>% 
  filter(state == "Florida") %>% 
  janitor::clean_names() %>% 
    mutate(binned_inc = as.factor(binned_inc)) %>% 
  mutate(binned_inc = fct_reorder(.f = binned_inc,.x = med_income)) 
can_filt_florida$top_can <- apply(can_filt_florida[,cancer_names],1,which.max) %>% map(.x = ., .f = ~ifelse(length(.)==0,"none",cancer_names[.])) %>% unlist() %>% as.factor()
#apply(can_filt[,cancer_names],1,which.max) %>% unlist() %>% table() %>% names(.) %>% as.numeric() %>% cancer_names[.]
can_filt_florida <- can_filt_florida %>% 
  #transmute(avg_household_size = log(avg_household_size)) %>% 
#  mutate()
#  .[c(-1009,-2221,-2090,-1008,-2087,-2218, -1007,-1006,-1005,-1004,-1003, -1002),] %>% 
dplyr::select( -pop_est2015,  -avg_ann_count, -avg_deaths_per_year, -median_age, -median_age_male,        -pct_no_hs18_24, -pct_hs18_24, -pct_bach_deg18_24, -pct_some_col18_24,  -percent_married, -pct_employed16_over, -pct_private_coverage, -pct_private_coverage_alone, -pct_emp_priv_coverage, -pct_public_coverage, -pct_black, -pct_asian, -pct_other_race, -state_abb,   -geography,  -binned_inc, -uterine, -bladder, -ovary, -brca_cis, -cervix, -colorectal, -melanoma, -ovary, -prostate, -brca, -study_per_cap)

```
```{r}
X = as.matrix(can_filt_florida %>% dplyr::select(-target_death_rate, -state, -county, -state_region) %>% sapply(as.numeric) %>% na.omit())
Y = na.omit(can_filt_florida)$target_death_rate
set.seed(1)
train<-sample(1:nrow(X),nrow(X)/2)
test<-(-train)
Y.test<-Y[test]
set.seed(2)
doParallel::registerDoParallel()
cv.out<-cv.glmnet(X[train,],Y[train], alpha=0.5,nfolds = 10, parallel = TRUE, lambda = grid)
plot(cv.out)
best.lambda<-cv.out$lambda.min
lasso.pred <- predict(cv.out,s=best.lambda,newx=X[test,])
final_lasso_florida_MSE <-mean((lasso.pred-Y.test)^2) #MSE
lasso3<-glmnet(X, Y, alpha=1, lambda=best.lambda)
res_lasso <- coef(lasso3) 
res_lasso %>% as.matrix() %>% as.data.frame() %>% add_rownames() %>% as.tibble() %>% 
  mutate(abs_lasso = abs(s0)) %>% 
  arrange(-abs_lasso) %>% knitr::kable()
final_lasso_florida<-lasso3
final_lasso_florida$dev.ratio
final_lasso_florida_range<-glmnet(X, Y, alpha=1, lambda=grid)
plotmo::plotres(final_lasso_range, info = T,which = 1:4) 
spacingTest(X,Y) %>% knitr::kable()
```
What's rather amazing here is that the proportion of deviance explained by this model explains a whopping `r final_lasso_florida$dev_ratio*100` % of the variance. The model looks reasonable, with the exception of two outliers. The target death rate there is roughly 3 times the national average, the poverty percent is above average, median income is below, and the percent of individuals with a college education hits the single digits. This is an outlier in the sense that it is a point that is apart from the rest in a res vs fitted plot, but it is a real feature of the data and should not be removed. Upon looking for evidence of htis, a recent article in JAMA cited this information (https://www.ncbi.nlm.nih.gov/pubmed/28118455), specifically mentioning this county last year. The article also mentions Utah and Idaho as low regions, just as we found.
```{r}

can_filt_florida[c(37,62),]
```
Two outliers are noted above. Throughout the data, there tend to be areas where the incidence of lung cancer is high and cancer death rate is high, which influences the model. Admittedly this may cause misleading results, but the data looks to be not in error and this sort of relationship was what was sought to be captured from the beginnning.

```{r, fig.width=10,fig.height=10,fig.align="center"}
#add in smoking, clinincal data, etc..
corrplot(cor(cbind(X,Y),method="kendall"),method = "number",order = "hclust")
corrplot(cor(cbind(X,Y),method="spearman")-cor(cbind(X,Y),method="pearson"),method = "number",order = "hclust")
corrplot(cor(cbind(X,Y)[,c("Y","incidence_rate","lungbronc")],method="kendall"),method = "number",order = "hclust")
corrplot(cor(cbind(X,Y)[,c("Y","incidence_rate","lungbronc")],method="pearson"),method = "number",order = "hclust")
```
It's actually quite remarkable... union county was mentioned in the JAMA article as well. The outlier also agrees with the data model, the poverty level is above the national average, the median income is low, the incidence rate is 3 times higher than that of the next highest county. Almost zero of the population possess a college education and the percent of public coverage is lower than the national average of `r mean(can_filt4$pct_public_coverage_alone)`%.
```{r}
ggplot(data=as.data.frame(cbind(X,Y)),aes(y=Y,x=lungbronc)) + geom_point()
can_filt_florida %>% 
  arrange(-lungbronc,-target_death_rate) %>% 
  select(county,state,lungbronc,target_death_rate,everything())

```

```{r}

```

Education seems to play a role in many sections of this data and it seems to be where the strongest signatures come from. Behavioral data should be added along with smoking rates. The BRFSS data is not subsampled, but just contains sparse responses. However, it is a currently used resource in literature precisely for this topic (https://pophealthmetrics.biomedcentral.com/articles/10.1186/1478-7954-12-5).
```{r eval=F} 
abb2state <- state.name
names(abb2state) <- toupper(state.abb)
# can_inc_by_coun_st <- can_inc_by_coun %>% 
#   mutate(state_abb = state2abb[state_name])

can_filt4_brfss_smoke <- p8105.datasets::brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(
    question=="Adults who are current smokers (variable calculated from one or more BRFSS questions)" & response=="No" & year==2010
#str_detect(string = Question,pattern = "smok")
) %>% 
  mutate(smoking_pct = 1-data_value) %>% 
  separate(locationdesc,into=c("state", "county"),sep = " - ") %>% 
  mutate(state = trimws(state),
         county = trimws(county)) %>% 
    mutate(state_name = abb2state[state]) %>% 
  select(state_name,everything()) %>% 
  inner_join(can_filt4, by=c("state_name" = "state", "county" = "county")) %>% 

  #convert state abb to state name
  select(state_name,county,smoking_pct, everything()) %>% 
  #select(locationabbr, locationdesc, data_value) %>% 
  arrange(-smoking_pct)
#stringsplit
#tolower
#inner join.
cor(can_filt4_brfss_smoke$smoking_pct,can_filt4_brfss_smoke$target_death_rate,use="pairwise.complete.obs")
cor(can_filt4_brfss_smoke$smoking_pct,can_filt4_brfss_smoke$lungbronc,use="pairwise.complete.obs")
lm(data = can_filt4_brfss_smoke, target_death_rate ~ smoking_pct) %>% broom::tidy() #%>% View
```
In the south region, we'll also take a look at e-cigarette utilization and cigarette utilization.
```{r smok_ecig, results = 'asis'}
smok_rate <- read_csv("./smoking/Behavioral_Risk_Factor_Data__Tobacco_Use__2011_to_present_.csv") %>% 
  janitor::clean_names() %>% 
  filter(year == "2017" & topic_desc == "Cigarette Use (Adults)" & race == "All Races" & response == "Never" & gender == "Overall") %>% 
  mutate(pct_have_smoked = 1-data_value) %>% 
  mutate(pct_never_smoker = data_value) %>%
  select(location_desc,data_value,pct_have_smoked,pct_never_smoker,location_abbr) %>% 
  arrange(-pct_have_smoked)
smok_rate %>% knitr::kable()
ecig_rate <- read_csv("./smoking/Behavioral_Risk_Factor_Data__Tobacco_Use__2011_to_present_.csv") %>%
  janitor::clean_names() %>% 
  filter(year == "2017" & topic_desc == "E-Cigarette Use (Adults)" & race == "All Races" & response == "Never" & gender == "Overall") %>% 
  mutate(pct_have_smoked = 1-data_value) %>% 
  mutate(pct_never_smoker = data_value) %>%
  select(location_desc,data_value,pct_have_smoked,pct_never_smoker,location_abbr) %>% 
  arrange(-pct_have_smoked)
#ecig_rate$year %>% unique()
ecig_rate %>% knitr::kable()
```

